\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-05-21 doc/paper.tex
edits up to Results

Same source analyses are a major part of a Forensic Toolmark Examiner's
job. In current practice, examiners make these comparisons by visual
inspection under a comparison microscope and come to one of the
following four conclusions: identification, inconclusive, elimination or
unsuitable for examination\textasciitilde{}\citep{afte-toolmarks1998}.
These conclusions are made on the basis of ``unique surface contours''
of the two toolmarks being in ``sufficient agreement''
\citep{afte-toolmarks1998}. AFTE describes the term ``sufficient
agreement'' as the possibility of another tool producing the markings
under comparison, as practically impossible \citep{afte-toolmarks1998}.
Potential subject bias in the assessment as well as the lack of
specified error rates are the main points of criticisms first raised by
the National Research Council in 2009 \citep{NAS:2009} and later
emphasized further by the President's Council of Advisors on Science and
Technology \citep{pcast2016}.

@@

Striation marks on bullets are made by impurities in the barrel. As the
bullet travels through the barrel, these imperfections leave
``scratches'' on the bullet surface. Typically, only striation marks in
the land engraved areas (LEAs) are considered \citet{afte-article1992}.
Bullet lands are depressed areas between the grooves made by the rifling
action of the barrel. Compared to toolmarks made by screwdrivers
striation marks on bullets are typically much smaller, both in length
and in width. Bullets also have a curved cross-sectional topography.
Figure \ref{fig:rgl} shows us how the signature from a bullet land
(bottom) lines up with the image of the land (top) from which it was
extracted. We can also see in the figure how the depth and relative
position of the striation markings seen in the image are interpreted as
the signature.

Bullet matching methods are usually based on these associated
signatures. \citet{chu2013} use an automatic method for counting
consecutive matching striae (CMS). The authors report an error rate of
52\% of the known same source lands comparisons as misidentified (false
negative) and zero false positives for known different source lands.
\citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation
function) and its discriminating power and applicability for same-source
analyses of bullets, but do not provide any error rates in their
discussion. \citet{aoas} use multiple features like CCF, CMS, D
(distance measure) etc in a random forest based method and compare every
land against every other land of digitised versions of Hamby 252 and
Hamby 44 \citep{hamby} published on the NIST Ballistics Database
\citep{nist}. The authors report an out-of-bag overall error rate of
0.46\%, comprised of an error rate of 30.05\% of same-source pairs that
were not identified and an error rate of 0.026\% of different-source
pairs that were incorrectly identified as same-source.

The Chumbley score provides us with another approach in the same-source
assessment of bullet striation marks. \citet{chumbley} compare two
toolmarks for same-source. The data for this study was obtained from 50
sequentially manufactured screwdriver tips. \citet{chumbley} report
error rates for markings made by the tips at different angles. For
markings made at 30 degree the authors report an average false negative
error rate of 0.089 and an average false positive error rate of 0.023.
For other angles of 60 and 85 degrees the false negatives error rate is
0.09 while the rate of false positives decreases to 0.01. The paper by
\citet{hadler} is based on the same data but the authors focus on
markings made under the same angle. The error rates associated with the
deterministic version of the score are 0.06 for false negatives and a
false positive error rate of 0.

@@

Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit or Gaussian filter. This effectively removes topographic
structure from the data in the attempt to increase the signal to noise
ratio. The span of the loess fit was found using cross-validation, as
described by \citet{aoas}.

There are two sources of scans for sets from the Hamby study available
to us: scans of Hamby 44 and Hamby 252 are available from the NIST
database \citep{nist}. Hamby 44 has also been made available to us and
has been scanned locally for CSAFE at the Roy J.~Carver High Resolution
Microscopy Facility using a Sensofar confocal light microscope. Scans in
the NIST database are made with a NanoFocus at 20x magnification. The
resolutions of the two instruments are different: the NIST scans are
taken at a resolution of 1.5625 \(\mu m\) per pixel, while the CSAFE
scans are available at a resolution of 0.645 \(\mu m\) per pixel. The
length of an average bullet land from Hamby (9 mm Ruger P85) is about 2
millimeter, resulting in signatures of about 1200 pixels for NIST scans,
and about 3000 pixels for CSAFE scans.

@@

The Chumbley score algorithm takes input in form of two digitized
toolmarks. The toolmark is in form of \(z(t)\) which is a spatial
process for location indexed by \(t\). \(t\) here denotes equally spaced
pixel locations for the striation marks under consideration,
\(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector of markings of
length \(s\) starting in location \(t\).

Let \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
be two digitized toolmarks (where \(T_1\) and \(T_2\) are not
necessarily equal). The toolmarks under consideration are potentially
from two different sources or the same source. \(T_1\) and \(T_2\), as
represented above, are the final pixel indexes of each marking and
therefore give the respective lengths of the markings.

In a pre-processing step the two markings are smoothed using a lowess
\citep{lowess} with coarseness parameter \(c\). Originally, this
smoothing is intended to remove drift and (sub)class characteristics
from individual markings, however, in the setting of matching bullet
striae, we can also make use of this mechanism to separate bullet
curvature in profiles from signatures before matching signatures.

After removing sub-class structure, the Chumbley scores is calculated in
two steps: an optimization step and a validation step. In the
optimization step, the two markings are aligned horizontally such that
within a pre-defined window of length \(w_o\) the correlation between
\(x(t_1)\) and \(y(t_2)\) is maximized: \[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
\] This results in an optimal vertical (in-phase) shift of
\(t_1^o - t_2^o\) for aligning the two markings.
{\textcolor{orange}{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$. }}

In the validation step, two sets of windows of size \(w_v\) are chosen
from both markings (see Figure \ref{fig:win-comparison}). In the first
set, pairs of windows are extracted from the two markings using the
optimal vertical shift as determined in the first step, whereas for the
second set the windows are extracted using a different (out-of-phase)
shift.

@@

Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
pairs \((s_i^{(1)}, s_i^{(2)})\) for all integer values \(i\) for which
both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
different-shift pairs are defined as \((s_i^{(1)}, s_{-i-1}^{(2)})\) for
all \(i\) where both \(s_i^{(1)}\) and \(s_{-i-1}^{(2)}\) are defined
(see \autoref{sketch-same-diff}).

@@

\caption{Two markings made by the same source. For convenience, the markings are moved into phase on the left and out-of phase on the right. In-phase (left) and out-of-phase (right) samples are shown by the light grey background. The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets of samples.}\label{fig:win-comparison}
\end{figure}

For both same- and different-shift pairs correlations between the
markings are calculated. The intuition here is that for two markings
from the same source the correlation for the in-phase sample should be
high, while the correlations of the out-of-phase sample provide a
measure for the base-level correlation for non-matching marks of a given
length \(w_v\). The Chumbley score is then computed as a Mann Whitney U
statistic to compare between in-phase sample and out-of-phase sample. In
the original method proposed in \citet{chumbley} both in-phase and
out-of-phase sample are extracted randomly, whereas \citet{hadler}
proposed the above specified deterministic rules for both samples to
make the resulting score deterministic while simultaneously avoiding
overlaps within selected marks to ensure independence.

@@

Looking closer at \autoref{eq.start}, we see that by definition, some
number of tests will fail to produce a result.
{\textcolor{orange}{The problem of failed tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the authors do not provide any percentage of how many tests failed for their data.}}
The algorithm fails to produce for two reasons: either the number of
eligible same-shift pairs is zero, or the number of different-shift
pairs is zero.

The number of same-shift pairs will be zero, if the optimal locations
\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
\(w_v\) are left on the same sides of the optimal locations, i.e.
\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\), i.e.~we have a
failure rate of \[
P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
\] While we can assume that once (sub-)class characteristics are
removed, optimal locations \(t_i^{o}\) are uniformly distributed across
the length of the profile, we cannot assume that \(t_1^o\) and \(t_2^o\)
are independent of each other. In particular, for same-source profiles,
we would expect a strong dependency between these locations, in which
case a large difference between locations is unlikely. However, for
different source matches, we can assume that locations are independent.
In that case, we expect a test to fail with a probability of
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
probability is about 0.0015.

The number of possible different-shift pairs also depends on the
location of the optimal locations \(t_1^o\) and \(t_2^o\). Whenever the
optimal locations are close to the boundaries, the number of possible
pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
\(t_i^{(o)} > T_i-w_o- w_v\). Assuming a correlation between optimal
locations \(t_1^o\) and \(t_2^o\) of close to one for same-source
profiles, this results in an expected rate of failure of
\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels. Assuming
independence in the optimal locations for different source profiles the
expected probability for a failed test is, again,
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\).

@@

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation} i.e.~for an odd number of same-shift correlations, we
skip the middle pair for the different-shift correlations (see
\autoref{sketch-diff-2}). This pairing ensures that the number of
different-shift pairings is the same or at most one less than the number
of same-shift pairings in all tests.
{\textcolor{orange}{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as{\bf (CS1)} and the suggested modified algorithm as{\bf (CS2)} and compare their performance on the available scans of the Hamby study.}}

@@

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}

\caption{\label{same-shift-failure}Sketch of same-shift pairings  (top) when the lag is too large to accomodate a a vaildation window in either of the two signatures}
\end{figure}

\caption{\label{diff-shift-failure}Sketch of diff-shift pairings  (top) when the number of diff-shift computations is likely to 0}
\end{figure}

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail}
\end{figure}

@@

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail using CS1. The highest percentage of failed tests under CS2
is 1.3\% for different source tests using a validation window size
\(w_v\) of 60 pixels.
{\textcolor{orange}{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans. }}{\textcolor{orange}{ Observed failure rates are higher than expected. This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform. }}

@@

{\textcolor{orange}{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3. The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified. 
The histograms in the bottom row show the relative optimal location $t^*$. Optimal locations are distributed uniformly once (sub-)class characteristics are removed. However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}}

@@

>>>>

Same-source analyses are a major part of a Forensic Toolmark Examiner's
job. In current practice, examiners make these comparisons by visual
inspection under a comparison microscope and come to one of the
following four conclusions: identification, inconclusive, elimination or
unsuitable for examination \citep{afte-toolmarks1998}. These conclusions
are made on the basis of ``unique surface contours'' of the two
toolmarks being in ``sufficient agreement'' \citep{afte-toolmarks1998}.
AFTE describes the term ``sufficient agreement'' as the possibility of
another tool producing the markings under comparison, as practically
impossible \citep{afte-toolmarks1998}. Potential subject bias in the
assessment as well as the lack of specified error rates are the main
points of criticisms first raised by the National Research Council in
2009 \citep{NAS:2009} and later emphasized further by the President's
Council of Advisors on Science and Technology \citep{pcast2016}.

@@

Striation marks on bullets are made by impurities in the barrel. As the
bullet travels through the barrel, these imperfections leave
``scratches'' on the bullet surface (see top of \autoref{fig:rgl}).
Typically, only striation marks in the land engraved areas (LEAs) are
considered \citep{afte-article1992}. Bullet lands are depressed areas
between the grooves made by the rifling action of the barrel. Compared
to toolmarks made by screwdrivers striation marks on bullets are
typically much smaller, both in length and in width. Bullets also have a
curved cross-sectional topography.

In same-source comparisons this curvature is usually removed using some
form of Gaussian filter \citep{ma2004} or non-parametric smoothing
\citep{aoas}. \citet{chu2013} use an automatic method for counting
consecutive matching striae (CMS). The authors report an error rate of
52\% for known same-source land comparisons to be (incorrectly)
identified as different-source (false negative) and zero false positives
for known different-source lands. \citet{ma2004} and
\citet{vorburger2011} discuss CCF (cross-correlation function) and its
discriminating power and applicability for same-source analyses of
bullets, but do not provide any error rates in their discussion.
\citet{aoas} use multiple features, such as CCF, CMS, D (distance
measure), etc. in a random forest based method and compare every land
against every other land of digitised versions of Hamby 252 and Hamby 44
\citep{hamby} published on the NIST Ballistics Database \citep{nist}.
The authors report an out-of-bag overall error rate of 0.46\%, comprised
of an error rate of 30.05\% of same-source pairs that were not
identified and an error rate of 0.026\% of different-source pairs that
were incorrectly identified as same-source.

The Chumbley score provides us with another approach in the same-source
assessment of bullet striation marks. \citet{chumbley} compare two
toolmarks for same-source. The data for this study was obtained from 50
sequentially manufactured screwdriver tips. \citet{chumbley} report
error rates for markings made by the tips at different angles. For
markings made at 30 degree the authors report an average false negative
error rate of 0.089 and an average false positive error rate of 0.023.
For marks made under angles of 60 and 85 degrees, respectively, the
false negatives error rate is 0.09 while the rate of false positives
decreases to 0.01. The paper by \citet{hadler} is based on the same data
but the authors focus on markings made under the same angle. The error
rates associated with the deterministic version of the score are
reported as 0.06 for false negatives and 0 for false positives.

@@

Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit or Gaussian filter. This effectively removes topographic
structure from the data in the attempt to increase the signal to noise
ratio. \autoref{fig:rgl} shows how the signature from a bullet land
(bottom) lines up with the image of the land (top) from which it was
extracted. We can see in the figure how the depth and relative position
of the striation markings seen in the image are interpreted as peaks and
valleys in the signature.

There are two sources of scans for sets from the Hamby study available
to us: scans of Hamby 44 and Hamby 252 are available from the NIST
database \citep{nist}. The physical Hamby 44 set has also been made
available to us and has been scanned locally for CSAFE at the Roy
J.~Carver High Resolution Microscopy Facility using a Sensofar confocal
light microscope. Scans in the NIST database are made with a NanoFocus
at 20x magnification. The resolutions of the two instruments are
different: the NIST scans are taken at a resolution of 1.5625 \(\mu m\)
per pixel, while the CSAFE scans are available at a resolution of 0.645
\(\mu m\) per pixel. The length of an average bullet land from Hamby (9
mm Ruger P85) is about 2 millimeter, resulting in signatures of about
1200 pixels for NIST scans, and about 3000 pixels for CSAFE scans.

@@

A digitized toolmark forms a spatial process \(z(t)\) with location
indexed by \(t\). \(t\) here denotes equally spaced pixel locations for
the striation marks under consideration. For a toolmark consisting of
\(t\) pixels, \(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector
of markings of length \(s\) starting in location \(t\).

The Chumbley score algorithm takes input in form of two digitized
toolmarks:

Let \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
be two digitized toolmarks (where \(T_1\) and \(T_2\), the lengths of
the two marks, are not necessarily equal). The toolmarks under
consideration are potentially from two different-sources or the
same-source.

In a pre-processing step the two markings are smoothed using a lowess
\citep{lowess} with coarseness parameter \(c\). Originally, this
smoothing is intended to remove drift and (sub)class characteristics
from individual markings, however, in the setting of matching bullet
striae, we can also make use of this mechanism to separate bullet
curvature in profiles from signatures before matching signatures.
\autoref{fig:sigs-profiles} shows an example of a bullet land profile
(left) and the corresponding signature (right).

After removing sub-class structure, the Chumbley score is calculated in
two steps: an optimization step and a validation step. In the
optimization step, the two markings are aligned horizontally such that
within a pre-defined window of length \(w_o\) the correlation between
\(x(t_1)\) and \(y(t_2)\) is maximized: \[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1-w_o, 1 \le t_2 \le T_2-w_o} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
\] This results in an optimal vertical (in-phase) shift of
\(t_1^o - t_2^o\) for aligning the two markings. We will denote the
relative optimal locations as \(t_1^*\) and \(t_2^*\), where
\(t_i^* = t_i^o/(T_i-w_o)\) for \(i=1,2\), such that
\(t_1^*, t_2^* \in [0,1]\). Once (sub-)class characteristics are
removed, the relative optimal locations should be distributed according
to a uniform distribution in \([0,1]\).

In the validation step, two sets of windows of size \(w_v\) are chosen
from both markings, \autoref[see][]{fig:win-comparison}. In the first
set, pairs of windows are extracted from the two markings using the
optimal vertical shift as determined in the first step, whereas for the
second set the windows are extracted using a different (out-of-phase)
shift.

@@

Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
pairs \((s_i^{(1)}, s_i^{(2)})\) with integer values \(i\) for which
both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
different-shift pairs are defined as \((s_i^{(1)}, s_{-i-1}^{(2)})\) for
all \(i\) where both \(s_i^{(1)}\) and \(s_{-i-1}^{(2)}\) are defined
(see \autoref{sketch-same-diff}).

@@

\caption{Two same-source markings. For convenience, the markings are moved into phase on the left and out-of phase on the right. In-phase (left) and out-of-phase (right) samples are shown by the light grey background. The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets of samples.}\label{fig:win-comparison}
\end{figure}

For both same- and different-shift pairs correlations between the
markings are calculated. The intuition here is that for two markings
from the same-source the correlation for the in-phase sample should be
high, while the correlations of the out-of-phase sample provide a
measure for the base-level correlation for non-matching marks of a given
length \(w_v\). The Chumbley score is then computed as a Mann Whitney U
statistic to compare between in-phase sample and out-of-phase sample. In
the original method proposed in \citet{chumbley} both in-phase and
out-of-phase sample are extracted randomly, whereas \citet{hadler}
proposed the above specified deterministic rules for both samples to
make the resulting score deterministic while simultaneously avoiding
overlaps within selected marks to ensure independence.

@@

Looking closer at \autoref{eq.start}, we see that by definition, some
number of tests will fail to produce a result. The problem of failed
tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the
authors do not provide any percentage of how many tests failed for their
data. The algorithm fails to produce for two reasons: either the number
of eligible same-shift pairs is zero, or the number of different-shift
pairs is zero.

The number of same-shift pairs will be zero, if the optimal locations
\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
\(w_v\) are left on the same sides of the optimal locations, i.e.
\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\), i.e.~we have a
failure rate of \[
P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
\] While we can assume that once (sub-)class characteristics are
removed, optimal locations \(t_i^{o}\) are uniformly distributed across
the length of the profile, we cannot assume that \(t_1^o\) and \(t_2^o\)
are independent of each other. In particular, for same-source profiles,
we would expect a strong dependency between these locations, in which
case a large difference between locations is unlikely. However, for
different-source matches, we can assume that locations are independent.
In that case, we expect a test to fail with a probability of
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
probability is about 0.0015.

The number of possible different-shift pairs also depends on the
location of the optimal locations \(t_1^o\) and \(t_2^o\). Whenever the
optimal locations are close to the boundaries, the number of possible
pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
\(t_i^{(o)} > T_i-w_o- w_v\). Assuming a correlation between optimal
locations \(t_1^o\) and \(t_2^o\) of close to one for same-source
profiles, this results in an expected rate of failure of
\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels. Assuming
independence in the optimal locations for different-source profiles the
expected probability for a failed test is, again,
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\).

@@

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation} i.e.~for an odd number of same-shift correlations, we
skip the middle pair for the different-shift correlations (see
\autoref{sketch-diff-2}). This pairing ensures that the number of
different-shift pairings is the same or at most one less than the number
of same-shift pairings in all tests. In the remainder of the paper, we
will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)}
and the suggested modified algorithm as \textbf{(CS2)} and compare their
performance on the available scans of the Hamby study.

@@

@@

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail using CS1. The highest percentage of failed tests under CS2
is 1.3\% for different-source tests using a validation window size
\(w_v\) of 60 pixels.
{\textcolor{orange}{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans. }}{\textcolor{orange}{ Observed failure rates are higher than expected. This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform. }}

@@

\autoref{fig:profile-sketch} gives an overview of the effect of
different coarseness parameters: from left to right, coarseness levels
\(c\) are varied in steps of 0.05 from 0.1 to 0.3. The top row shows
resulting signatures after smoothing the profile shown in
Figure\textasciitilde{}\ref{fig:sigs-profiles} with different levels of
coarseness. The histograms in the bottom row show the relative optimal
location \(t^*\). Optimal locations are distributed uniformly once
(sub-)class characteristics are removed. However, for coarseness values
of \(c > 0.20\) we see quite distinct boundary effects: optimal
locations \(t^*\) are found at the very extreme ends of a profile more
often than one would expect based on a uniform distribution.

@@

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\begin{appendix}

\section{Scenarios of failed tests}

Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned. These failures are inherent to the Chumbley Score and cannot be prevented.
\begin{figure}[hbtp]
\centering
\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}

\caption{\label{same-shift-failure}Sketch of same-shift pairings. When the lag between optimal reference points is too large to accommodate a validation window in noth signatures, both CS1 and CS2 fail.}
\end{figure}

\autoref{diff-shift-failure} shows two situations, in which CS1 fails to identify any different-shift pairings. This happens, when both relative locations are close to either end of the signature. CS2 can still be computed in this situation as long as there are at least two same-shift pairs identified.

\caption{\label{diff-shift-failure}Sketch of different-shift pairings. For CS1 no different-shift pairs can be identified, resulting in a failed test.}
\end{figure}

<!--
\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail.}
\end{figure}
-->

\end{appendix}
